---
title: "430 Homework 3"
author: "Yinuo Song"
date: "11/24/2019"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction

The Retail Trade sector comprises establishments engaged in retailing merchandise, generally without transformation, and rendering services incidental to the sale of merchandise. The 
retailing process is the final step in the distribution of merchandise; retailers are, therefore, organized to sell merchandise in small quantities to the general public.  

I get my data source for retail trade from the U.S. Census Bureau website. It contains monthly sales for retail trade and food services, adjusted and unadjusted for seasonal variations within the U.S. area and has 67 columns in total. In my R project research, I compiled the data from 2000 to 2019 and I only use the total retail sales columns as my main factor. I plan to make the forecasting for the U.S. monthly retail sales in total for the next two years, using functions about time series in R and building models to make trend and seasonality forecast.

The data source is: https://www.census.gov/retail/index.html

```{r cars}
rm(list=ls(all=TRUE))
getwd()
library(lattice)
library(foreign)
library(MASS)
library(car)
require(stats)
require(stats4)
library(KernSmooth)
library(fastICA)
library(cluster)
library(leaps)
library(mgcv)
library(rpart)
library(pan)
library(mgcv)
library(DAAG)
library("TTR")
library(tis)
require("datasets")
require(graphics)
library("forecast")
require(astsa)
library(RColorBrewer)
library(plotrix)
library(xtable)
library(stats)
library(pastecs)
library(psych)
library(TSA)
library(timeSeries)
library(fUnitRoots)
library(fBasics)
library(tseries)
library(timsac)
library(TTR)
library(fpp)
library(RColorBrewer)
library(plotrix)
library(nlstools)
library(Metrics)
```

Read in the data

```{r}
data1 <- read.csv("US Census Bureau Data.csv")
head(data1)
retail <- data1[, 2]
retail_ts <- ts(retail, start=2000, freq=12)
describe(retail_ts)
t1 <- seq(2000, 2019.9, length=length(retail_ts))
```

Time-series plot

```{r}
plot(retail_ts, ylab="Retail sales", xlab="Time", lwd=2, col='skyblue3')
```

We can see from the plot that it does not suggest that the data are covariance stationary. 
Covariance stationarity refers to a set of conditions on the underlying probability structure of a time series that has proven to be very especially valuable in this regard. A time series yt is said to be covariance stationary if it meets the following conditions: 1.Constant mean 2.Constant (and finite) variance 3.Stable autocovariance function. My data does not have these futures, so the plot are not covariance stationary.

ACF & PACF

```{r}
quartz()
par(mfrow=c(3,1))
acf(retail_ts, type = "covariance", main="Autocovariance", lag.max=50, ylab="COV")
acf(retail_ts, type = "correlation", main="Autocorrelation", lag.max=50, ylab="ACF")
acf(retail_ts, type = "partial", main="Partial Autocorrelation", lag.max=50, ylab="PACF")
```

From the ACF and PACF test, there is dependence on how retail sales have changed over time. 

```{r}
#White noise test
acf_val=acf(retail_ts)$acf
Box.test(acf_val, type = "Ljung-Box")
Box.test(acf_val, type = "Box-Pierce")
```

```{r}
#MA(q) Model Fitting
ma1=arma(retail_ts, order=c(0,1)) 
summary(ma1)
ma2=arma(retail_ts, order=c(0,2)) 
summary(ma2)
ma3=arma(retail_ts, order=c(0,10))
summary(ma3)
```

```{r}
#All the MA(q) fits
quartz()
plot(retail_ts, xlab='Time', ylab="Retail sales", lwd=2)
grid()
lines(ma1$fitted.values, col="blue", lwd=2)
lines(ma2$fitted.values, col="seagreen2", lwd=2)
lines(ma3$fitted.values, col="red", lwd=2)
legend("topright", legend=c("Data","MA(10)","MA(2)","MA(1)"), text.col=1:4, bty="n")
```

```{r}
#Examine the best fit MA(q) model
quartz()
par(mfrow=c(2,2))
plot(retail_ts, xlab='Time', ylab="Retail sales", lwd=2)
lines(ma3$fitted.values, col="red", lwd=1, lty=1)
plot(ma3$residuals, ylab="Residuals")
acf(ma3$residuals[12:136], type = "correlation", main="Autocorrelation",lag.max=13,ylab="ACF")
acf(ma3$residuals[12:136], type = "partial",main="Partial Autocorrelation",lag.max=13, ylab="PACF")
```

We can see a significant improvement from looking at the ACF and PACF plots, the residuals 
now look consistent with noise, suggesting we accounted for most of the dynamics left after 
detrending and seasonally adjusting the data.

Trend fitting

```{r}
#Mode1: Linear Fit model 
m1 <- lm(retail_ts ~ t1)

#Mode2: Log-Linear Fit model 
m2 <- lm(log(retail_ts) ~ t1) 

#Mode3: Log-quadratic model
lretail <- log(retail_ts)
m3 <- lm(lretail ~ t1+t1^2)
```

Figures

```{r}
plot(retail_ts, ylab="Retail sales", xlab="Time", lwd=2, col='skyblue3')
lines(t1, m1$fit, col="red3", lwd=2)

plot(lretail, ylab="Retail sales", xlab="Time", lwd=2, col='skyblue3', main="Log-Linear Fit")
lines(t1, m2$fit, col="blue", lwd=2)

plot(lretail, ylab="Retail sales", xlab="Time", lwd=2, col='skyblue3', main="Log-Linear Fit")
lines(t1, m3$fit, col="darkblue", lwd=2)
```

From the plot, we can see that the three models seems to be similar with each other.

Residuals & Fitted values 

```{r}
#Mode1: Linear Fit model
plot(t1, m1$res, ylab="Residuals", type='l', xlab="Time", main="Linear residuals")
abline(h=0, col="red", lwd=2)

#Mode2: Log-Linear Fit model 
plot(t1, m2$res, ylab="Residuals", type='l', xlab="Time", main="Log-Linear residuals")
abline(h=0, col="red", lwd=2)

#Mode3: Log-quadratic model
plot(t1, m3$res, ylab="Residuals", type='l', xlab="Time", main="Log-quadratic residuals")
abline(h=0, col="red", lwd=2)
```

I observed that the residuals are showing the periodic characteristics, and the residuals 
around 2008 year are much larger and unstable when compared with other years. This may be 
because of the financial crisis at that time.

Histogram of residuals

```{r}
par(mfrow=c(3,1))
hist(m1$res, breaks = "FD", col = "skyblue2", freq = FALSE, ylab = "Density")
lines(density(m1$res), lwd = 2, col = "red")

hist(m2$res, breaks = "FD", col = "blue", freq = FALSE, ylab = "Density")
lines(density(m2$res), lwd = 2, col = "red")

hist(m3$res, breaks = "FD", col = "darkblue", freq = FALSE, ylab = "Density")
lines(density(m3$res), lwd = 2, col = "red")
```

I observed that the values of the density for the residuals are different from each model,
but all the residuals from the six models are close to the normal distribution.

Diagnostic statistics

```{r}
m1 <- lm(retail_ts ~ t1)
summary(m1)

m2 <- lm(log(retail_ts) ~ t1)
summary(m2)

m3 <- lm(lretail ~ t1+t1^2)
summary(m3)
```

R squared of the linear model is 0.8207, R squared of the log-linear model is 0.8278, 
R squared of the Log-quadratic model is 0.8278. It suggests that these three models are 
all good fit models. And the model 2 and model 3 are better than model 1. P-values are 
all small but far bigger than the t value, which suggests that the estimators are 
statistically significant. 

AIC & BIC

```{r}
AIC(m1,m2,m3)
BIC(m1,m2,m3)
```

From the AIC and BIC tests, we can tell that the model 2 and model 3 are better than the
model 1 since their test values are smaller. Model 2 and model 3 have same AIC and BIC values, so either model is great, and I plan to choose the model 3 as my preffered model.

Trend forecasting 

```{r}
m3 <- lm(lretail ~ t1+t1^2)
tn=data.frame(t1=seq(2019.9, 2021.9, length=24))
pred.m3=predict(m3, tn, se.fit = TRUE)
pred.plim.m3=predict(m3, tn, n.ahead=60, level = 0.95, interval="prediction")
pred.clim.m3=predict(m3, tn, n.ahead=60, level = 0.95, interval="confidence")
matplot(tn$t1, cbind(pred.clim.m3, pred.plim.m3[,-1]),
        lty=c(1,1,1,3,3), type="l", lwd=2, ylab="predicted y",xlab="Time")
```

I predict the US monthly retail sales for the next two years as the plot shows. We can see
the prediction interval for different confidence levels in this plot. It shows clearly that
the US monthly retail sales will increases in the next two years.

```{r}
#Holt-Winters Filter
HoltWinters(retail_ts)
quartz()
plot(retail_ts, xlab="Time", ylab="Retail sales")
lines(HoltWinters(retail_ts)$fitted[,1], col="red")

#Holt-Winters Prediction
retail_ts.hw <- HoltWinters(retail_ts)
predict(retail_ts.hw, n.ahead=12)
quartz()
plot(retail_ts, xlim=c(2000,2039), xlab="Year", ylab="Retail sales", ylim=c(200000,550000))
lines(predict(retail_ts.hw, n.ahead=48), col=2)
```

Trend Forecasting 

```{r}
quartz()
par(mfrow=c(2,1))
plot(retail_ts, main="Data", xlab="Time", ylab="Retail sales")
plot(forecast(retail_ts), main="Data with Respective Point and Interval Forecasts",
     xlab="Time", ylab="Retail", shadecols="oldstyle")
```

Model with a full set of seasonal dummies

```{r}
m4 <- tslm(retail_ts ~ season)
summary(m4)
```

We can see that the p-values of most seasons are very small which suggest statistically 
significance. Therefore, we can say that there are seasonal factors in this model. 

Plot seasonal factors

```{r}
plot(m4$coef,type='l',ylab='Seasonal Factors', xlab="Season",lwd=2, main="Plot of Seasonal Factors")
```

We can see that the seasonal factors are extremely significant when approaching Christmas
and New Years Day. It fits well with the fact that people will buy much more things during 
these holidays. The seasonal factors drop very low after these holidays. 

Full model

```{r}
m5 <- tslm(lretail ~ t1+t1^2+season)
par(mfrow=c(3,1))
plot(lretail, xlab="Time", ylab="Retail sales")
lines(t1,m5$fit,col="red3",lwd=2)
plot(t1, m5$res, ylab="Residuals", type='l', xlab="Time")
hist(m5$res, main="Histogram of Residuals", col="skyblue3")
```

Summary statistics and the error metrics of full model

```{r}
S(m5)
```

In this full model, we can see that except for season 2(Feb), all other months have strong 
significant seasonal effects on the retail sales. In details, people tend to buy more goods
in April, September and October. And they will buy fewer goods during other months.

```{r}
mape(lretail, m5$fitted.values)
mse(lretail, m5$fitted.values)
bias(lretail, m5$fitted.values)
mae(lretail, m5$fitted.values)
```

MAPE: The caculated value of MAPE is 0.0025, which is close to zero.
MSE: The caculated value of MSE is 0.0018, which is close to zero.
BIAS: The caculated value of BIAS is so small and is close to zero.
MAE: The caculated value of MAE is 0.03 which is very small.
Therefore, my full model m5 is great.

Seasonality Forecasting

```{r}
m5 <- tslm(lretail ~ t1+t1^2+season)
tn=data.frame(t1=seq(2019.9, 2021.9, length=length(lretail)))
pred.m5=predict(m5, tn, se.fit = TRUE)
pred.plim.m5=predict(m5, tn, n.ahead=60, level = 0.95, interval="prediction")
pred.clim.m5=predict(m5, tn, n.ahead=60, level = 0.95, interval="confidence")
matplot(tn$t1, cbind(pred.clim.m5, pred.plim.m5[,-1]),
        lty=c(1,1,1,3,3), type="l", lwd=2, ylab="predicted y",xlab="Time")
```

I predict that the next two years as the graph shows. We can see that the US retail sales
will follow the seasonal regulation with an increasing trend.

STL

```{r}
plot(stl(lretail,s.window="periodic"))
summary(forecast(lretail))
```

From the STL graph, we can see that the seasonal effect and long-term trend are independent.
Therefore, I choose the additive model.

Seasonal adjusted plot 

```{r}
decompose_lretail <- decompose(lretail, "additive")
adjust_lretail <- lretail-decompose_lretail$seasonal
plot(adjust_lretail)
```

After the seasonal adjustment, I think my model is ok. From the graph, we can see an
increasing trend of the data with seasonal trend.

Improve the model through ets

```{r}
plot(lretail,s.window="periodic")
forecast(lretail)
plot(forecast(lretail))

fit=ets(lretail)
plot(fit)
plot(forecast(fit,level=c(50,80,95),h=24))
```

We can see that the seasonal effect is linked to time except for the year around 2008,
it may be because of the financial crisis happened at that year.

Conclusion
From the analysis above, we can conclude that the US monthly retail sale has a trend that 
increases as the time goes by except for the time when economics crisis happens. Meantime,
it has very significant seasonal effects. That is to say, people tend to buy more goods in
specific months such as April, September and October, and fewer in other months. And the 
seasonal effects are stable as the time goes by. 

Future work
I can try to analyze the specific time around 2008 to make the model more perfect by using 
the sectional functions. I believe that my model will become even better than this model in
that way 

Reference:
https://www.census.gov/retail/index.html




